### 2022-01-09 paper/manuscript.Rmd
update standardization

Penalized regression analyses often require the scales of the predictors to be equalized ([@tibshirani1996; @lee2015note]). This is because $\lambda$, the regularization parameter, penalizes the magnitude of the coefficients that are dependent on the scales of the predictors. This leads to the unfair penalization of larger coefficients that do not necessarily reflect importance ([@zou2006adaptive]). For example, in the case of 2 or more equally important predictors, the coefficient of the predictor with the largest scale will be shrunk most.

To equalize predictor scales, an often-used technique is standardization ([@gelman2008scaling]). Standardization sets the mean of all predictors to 0 and their variance to 1. It does so by transforming all observations in the predictor space $\boldsymbol{X}$ to Z-scores as follows:

@@

There is some debate regarding the standardization of binary/dummy predictors in penalized regression ([@lee2015note]). Some suggest to always standardize dummy variables ([@tibshirani1997lasso]). However, standardizing binary predictors increases the absolute maximum values the predictor can take on which in some cases can make it more sensitive to the penalty function ([@harrell2001regression]).

>>>>

Penalized regression analyses often require the scales of the predictors to be equalized ([@tibshirani1996; @lee2015note]). This is because $\lambda$, the regularization parameter, penalizes the magnitude of the coefficients that are dependent on the scales of the predictors. This can lead to the unfair penalization of larger coefficients that do not necessarily reflect importance ([@zou2006adaptive]). To clarify, a regression parameter $\beta$ can be interpreted as the amount outcome $Y$ increases as predictor $X$ increases with 1 unit. This means that as $X$ is reduced by a factor of 10, its regression coefficient is increased by a factor of 10. So suppose the desire to predict running speed by a person`s height. If height is measured in kilometers, than a one unit increase in height, which is a one kilometer increase, will likely lead to a drastic increase in running speed. In penalized regression techniques, this high value for $\beta$ would then dominate the cost function, leaving coefficients of other predictors unregularized.

To equalize predictor scales, an often-used technique is standardization ([@gelman2008scaling]). Standardization sets the mean of all predictors to 0 and their standard deviation to 1. It does so by transforming all observations in the predictor space $\boldsymbol{X}$ to Z-scores as follows:

@@

There is some debate regarding the standardization of binary/dummy predictors in penalized regression ([@lee2015note]). Some suggest to always standardize dummy variables ([@tibshirani1997lasso]). However, standardizing binary predictors increases the absolute maximum values the predictor can take on which in some cases can make its coefficient more sensitive to the penalty function ([@harrell2001regression]).
